---
title: "Problem Set 1"
author: "Alex, Daniel and Micah"
date: "8/27/2019"
output: github_document
---

```{r setup, include=FALSE}
library(data.table)
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
```

# Potential Outcomes Notation 

1. Explain the notation $Y_i(1)$.  
**Answer:** This notation indicates the potential treatment 1 for i treatment groups.  
2. Explain the notation $Y_1(1)$.  
**Answer:** This notation indicates the potential treatment 1 for group 1 of i. 
3. Explain the notation $E[Y_i(1)|d_i=0]$.  
**Answer:** This notation means the expectation of Y_i(1) when one group is selected randomly from those groups which were not treated.  
4. Explain the difference between the notation $E[Y_i(1)]$ and $E[Y_i(1)|d_i=1]$.  
**Answer:** E[Y_i(1)] means when one subject from the treatement group is selected at random. E[Y_i(1)|d_i=1] is when one treated subject from the treatement group is selected at random.  The former group may select a subject in the treatment group which was not treated.  

# Potential Outcomes and Treatment Effects 

1. Use the values in the table below to illustrate that $E[Y_i(1)]-E[Y_i(0)] = E[Y_i(1)- [Y_i(0)]$.  
**Answer:** Expected values still follow the associative property.  
```{r, include = TRUE}
y_0_ex <- c(10, 12, 15, 11, 10, 17, 16)
y_1_ex <- c(12, 12, 18, 14, 15, 18, 16)
tau_ex <- c(2, 0, 3, 3, 5, 1, 0)
tau.mean <- mean(tau_ex)
y_0_ex.mean <- mean(y_0_ex)
y_1_ex.mean <- mean(y_1_ex)
y.diff <- y_1_ex.mean - y_0_ex.mean
```
```{r}
cat("E[Y_i(1)- [Y_i(0)] = ",tau.mean)
cat("E[Y_i(1)]-E[Y_i(0)] = ",y.diff )

```

2. Is it possible to collect all necessary values and construct a table like the one below in real life? Explain why or why not.  
**Answer:** No, it is not possible to construct a table like the below. In the table below each subject is receiving both the treatment and the control conditions. This is not how a real field experiment's data would look.  

```{r, include = FALSE}
table <- data.table(
  subject = 1:7, 
  y_0 = c(10, 12, 15, 11, 10, 17, 16), 
  y_1 = c(12, 12, 18, 14, 15, 18, 16),
  tau = c(2, 0, 3, 3, 5, 1, 0)
)
```

```{r}
kable(table)
```

# Visual Acuity 

Suppose we are interested in the hypothesis that children playing outside leads them to have better eyesight.

Consider the following population of ten representative children whose visual acuity we can measure. (Visual acuity is the decimal version of the fraction given as output in standard eye exams. Someone with 20/20 vision has acuity 1.0, while someone with 20/40 vision has acuity 0.5. Numbers greater than 1.0 are possible for people with better than “normal” visual acuity.)

```{r, include = FALSE}
d <- data.table(
  child = 1:10, 
  y_0 = c(1.2, 0.1, 0.5, 0.8, 1.5, 2.0, 1.3, 0.7, 1.1, 1.4), 
  y_1 = c(1.2, 0.7, 0.5, 0.8, 0.6, 2.0, 1.3, 0.7, 1.1, 1.4)
)
```

```{r}
kable(d)
```

In this table, `y_1` means means the measured *visual acuity* if the child were to play outside at least 10 hours per week from ages 3 to 6. `y_0` means the measured *visual acuity* if the child were to play outside fewer than 10 hours per week from age 3 to age 6. Both of these potential outcomes *at the child level* would be measured at the same time, when the child is 6. 

1. Compute the individual treatment effect for each of the ten children.  
**Answer:**. 
```{r, include = TRUE}

y_0 <- c(1.2, 0.1, 0.5, 0.8, 1.5, 2.0, 1.3, 0.7, 1.1, 1.4)
y_1 <- c(1.2, 0.7, 0.5, 0.8, 0.6, 2.0, 1.3, 0.7, 1.1, 1.4)
tau <- y_1 - y_0

a1 <- data.table(
  child = 1:10, 
  y_0,
  y_1,
  tau
)
```

```{r}
kable(a1)
```

2. Tell a "story" that could explain this distribution of treatment effects. In particular, discuss what might cause some children to have different treatment effects than others.  
**Answer:** Assuming there is a normal distribution it makes sense that there is some variation within a sample.  However, to tell a "story" we can assume that a child with a decrease in *visual accuity* may have a genetic predisposition to poor eyesight. Often eyesight changes with age. This could be a larger influence than playing outside. In the case of a child with increased *visual acuity* perhaps the measurement of original *visual acuity* has an error. This is common within experimentation as accuracy cannot be guaranteed with measuring devices.  For instance, a child given a standard vision test may be able to guess some answers correctly when believing they will be praised or rewarded for doing well.  

3. For this population, what is the true average treatment effect (ATE) of playing outside.  
```{r}
tau.mean <- mean(tau)

print(tau.mean)
```

4. Suppose we are able to do an experiment in which we can control the amount of time that these children play outside for three years. We happen to randomly assign the odd-numbered children to treatment and the even-numbered children to control. What is the estimate of the ATE you would reach under this assignment? (Please describe your work.)  
**Answer:**  Sorry running out of time to complete otherwise I would have run loops and use modulo to select the correct position into a new list.  

First, the control group will only get readings on even numbered children and the treatment group will only have readings on the odd numbers.  We can use the associative properties from the previous question to calculate the mean on each group separately and find the different tau for the whole.
```{r}
y_0_even <- c(NULL, 0.1, NULL, 0.8, NULL, 2.0, NULL, 0.7, NULL, 1.4)
y_1_odd <- c(1.2, NULL, 0.5, NULL, 0.6, NULL, 1.3, NULL, 1.1, NULL)

y_0_even.mean <- mean(y_0_even)
y_1_odd.mean <- mean(y_1_odd)
tau.4 <- y_1_odd.mean - y_0_even.mean

print(tau.4)


```

5. How different is the estimate from the truth? Intuitively, why is there a difference?  
**Answer:** The estimate is likely `r tau.4 - tau.mean` away from the truth.  Intuitively the difference exists because of sample size where simple randomization caused error.  
6. We just considered one way (odd-even) an experiment might split the children. How many different ways (every possible ways) are there to split the children into a treatment versus a control group (assuming at least one person is always in the treatment group and at least one person is always in the control group)?  
**Answer:** This method assumes we are randomly assigning to one group and the remainder go to the other group. Technically I could divide this by 2 as the children would be randomly sorted the same.  
```{r}
answer <-{
  factorial(10)/(factorial(1)*factorial(9)) + 
    factorial(10)/(factorial(2)*factorial(8)) +
    factorial(10)/(factorial(3)*factorial(7)) +
    factorial(10)/(factorial(4)*factorial(6)) +
    factorial(10)/(factorial(5)*factorial(5)) +
    factorial(10)/(factorial(6)*factorial(4)) +
    factorial(10)/(factorial(7)*factorial(3)) +
    factorial(10)/(factorial(8)*factorial(2)) +
    factorial(10)/(factorial(9)*factorial(1)) 
}

print(answer)
```
7. Suppose that we decide it is too hard to control the behavior of the children, so we do an observational study instead. Children 1-5 choose to play an average of more than 10 hours per week from age 3 to age 6, while Children 6-10 play less than 10 hours per week. Compute the difference in means from the resulting observational data.  
**Answer:**  
```{r}
y_0_h1 <- c(NULL, NULL, NULL, NULL, NULL, 2.0, 1.3, 0.7, 1.1, 1.4)
y_1_h2 <- c(1.2, 0.7, 0.5, 0.8, 0.6, NULL, NULL, NULL, NULL, NULL)

y_0_h1.mean <- mean(y_0_h1)
y_1_h2.mean <- mean(y_1_h2)
tau.7 <- y_1_h2.mean - y_0_h1.mean

print(y_0_h1.mean)
print(y_1_h2.mean)
print(tau.7)


```
8. Compare your answer in (g) to the true ATE. Intuitively, what causes the difference?  
**Answer:** The answer this time is `r tau.7 - tau.mean` (-0.51) away from the true ATE. Again, sample size is causing the issue. Randomization is meant to equalize these effects, but with a sample size of n = 10 the variability is going to be large.  

# Randomization and Experiments 

1. Assume that researcher takes a random sample of elementary school children and compare the grades of those who were previously enrolled in an early childhood education program with the grades of those who were not enrolled in such a program. Is this an experiment or an observational study? Explain!  
**Answer:** This is an observational study. This could have been an experiment if a group of children were randomly assigned to take early childhood education or not. In the case described the randomization is within the population that was already presorted into two categories in a non-random fashion.  
2. Assume that the researcher works together with an organization that provides early childhood education and offer free programs to certain children. However, which children that received this offer was not randomly selected by the researcher but rather chosen by the local government. (Assume that the government did not use random assignment but instead gives the offer to students who are deemed to need it the most) The research follows up a couple of years later by comparing the elementary school grades of students offered free early childhood education to those who were not. Is this an experiment or an observational study? Explain!  
**Answer:** as above this method is not an experiment.  The assignment was not done randomly, but with a bias to those that are *deemed to need it the most*. This segmentation of the population may have significant biases.  
3. Does your answer to part (2) change if we instead assume that the government assigned students to treatment and control by "coin toss" for each student?  
**Answer:** Yes, this would essentially create a natural experiment. We can equate the *coin toss* to a lottery system. Factors like parental interest, race, ethnicity, economic status should be equally distributed based on the random assignment between treatment and control.  

# Moral Panic 

Suppose that a researcher finds that high school students who listen to death metal music at least once per week are more likely to perform badly on standardized test. As a consequence, the researcher writes an opinion piece in which she recommends parents to keep their kids away from "dangerous, satanic music". Let $Y_i(0)$ be each student's test score when listening to death metal at least one time per week. Let $Y_i(1)$ be the test score when listening to death metal less than one time per week. 

1. Explain the statement $E[Y_i(0)|D_i=0] = E[Y_i(0)|D_i=1]$ in words. First, state the rote english language translation; but then, second, tell us the *meaning* of this statement.  
**Answer:** The statement $E[Y_i(0)|D_i=0] = E[Y_i(0)|D_i=1]$ can be translated to the expected outcome of students' test scores when listening to death metal at least once per week and is not treated is equal to the expected outcome of students' test scores when listening to death metal at least once per week and is treated.  
The meaning of this is that students in the control group (listens to death metal once per week) have the same expected outcome whether they receive the treatment (listening to death metal less than once per week) or not.  
2. Do you expect the above condition to hold in this case? Explain why or why not.  
**Answer:** Yes, I expect the above condition to hold in this case.  It's highly unlikely that listening to death metal has a statistically significant causal relation to standardized test scores. More likely, is that this was a proxy variable for other confounding factors like economic status or academic interest.  

# MIDS Admission 

Suppose a researcher at UC Berkeley wants to test the effect of taking the MIDS program on future wages. The researcher convinces the School of Information to make admission into the MIDS program random among those who apply. The idea is that since admission is random, it is now possible to later obtain an unbiased estimate of the effect by comparing wages of those who where admitted to a random sample of people who did not take the MIDS program. Do you believe this experimental design would give you an unbiased estimate? Explain why or why not.  Assume that everybody who gets offer takes it and that prospective students do not know admission is random.  
**Answer:** It is not possible to obtain an unbiased estimate of the effect by comparing wages of those were admitted to those who were not admitted. In this experimental design those excluded from admission do not know admission was made randomly. This information being kept secret may then depress the wages of rejected candidates. They may take that as a market signal to not pursue opportunities which they otherwise would have. In this scenario the experimental design may actually widen the gap and increase the wage disparity of future earnings.  